SHIV: Reducing Supervisor Burden using Support Vectors to Efficiently Learn Robot Control Policies from Demonstrations [v17]

### 2015-09-02 paper.tex
Anca's comments

\newboolean{include-notes}
\setboolean{include-notes}{true}
\newcommand{\adnote}[1]{\ifthenelse{\boolean{include-notes}}%
 {\textcolor{green}{\textbf{AD: #1}}}{}}

@@

\author{Michael Laskey$^1$, Sam Staszak$^3$, Wesley Yu-Shu Hsieh$^1$, Jeff Mahler$^1$, Florian T. Pokorny$^1$, Anca Dragan$^1$, Ken Goldberg$^{1,2}$% <-this % stops a space
\thanks{$^1$ Department of Electrical Engineering and Computer Sciences; {\small \{mdlaskey,iamwesleyhsieh, ftpokorny, anca\}@berkeley.edu}}%
\thanks{$^2$ Department of Industrial Engineering and Operations Research; {\small goldberg@berkeley.edu}}%
\thanks{$^{1-2}$ University of California, Berkeley;  Berkeley, CA 94720, USA}%
}
\begin{document}

@@

\begin{abstract}
Online algorithms for robot model-free learning from demonstration (such as DAgger) can learn policies for problems where system
dynamics and cost function are unknown. However, during learning they impose a burden on supervisors to provide labels (control
signals) as the robot encounters new states when iteratively executing its current best policy. We introduce the SHIV
algorithm (Svm-based reduction in Human InterVention), which reduces this burden by enabling the robot to only request supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary high dimensional state distributions.  To facilitate scaling and outlier rejection, risk is based on distance to an approximate level set boundary defined by a One Class support vector machine.  We compare SHIV with DAgger in three policy learning experiments: 1) a driving simulator with a 27936D visual HOG feature state representation, 2) 4DOF robot arm push-grasping in clutter (simulation with Box2D) with a 22D state representation  of joint angles and the pose of each object, and 3) surgical needle insertion with a 16D state represented spatial transform.  Results suggest that SHIV can reduce the number of queries up to 70$\%$.

@@

In model-free robot learning from demonstration, a robot learns to perform a task, such as driving or grasping an object in a cluttered environment as shown in Fig. \ref{fig:teaser}, from examples provided by a competent supervisor, usually a human. The problem is model-free if  the system dynamics are unknown, eg. dynamics model may be highly non-linear or noisy. In learning from demonstration, the cost function, which is the negative of the reward function, is usually unknown or significantly delayed. In the grasping for clutter example, shown in Fig. \ref{fig:teaser}(b) the robot is only penalized when an object is knocked off the table, which means a potentially large number of states can be visited before any feedback. To potentially avoid this a supervisor provides guidance through demonstrations and the robot learns a policy mapping states to controls~\cite{argall2009survey}.

In the supervised setting, a policy is learned offline from a fixed set of examples. When the robot executes this policy, the sequential process can quickly evolve away from the previously explored set of states.  The learned policy might not be helpful in unfamiliar states, leading to unrecoverable failures.  For example, a robot driving may be trained on examples driving safely down the center of a lane, but even slight deviations will put the robot into states near the side of the road where it has not encountered before, where its policy could  fail~\cite{pomerleau1989alvinn}. It has been shown, the number of errors made by the robot in such cases  can scale quadratically with the time horizon of the task~\cite{ross2010efficient}.

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/teaser.pdf}
\caption{ 
\todo{update figure}
SHIV reduces the number of queries to the supervisor in our three test domains: 1) a driving simulator where the goal is to learn a controller on HOG features extracted from synthetic images to keep a car on a polygonal track; 2) push-grasping in clutter in a physics simulator, where the goal is to learn a controller from human demonstrations to grasp the green object while not letting the red square objects fall off the gray boundary representing a table; 3) learning a controller for the Da Vinci Research Kit from a master surgeon's demonstrations of correcting a needle before suturing in surgery. 
}
\vspace*{-10pt}
\label{fig:teaser}
\end{figure}

On-line learning from demonstration approaches such as DAgger iteratively gather more examples from the supervisor~\cite{grollman2007dogged,ross2010efficient,ross2010reduction}. In DAgger, a series of policies are learned.  In each iteration, the latest policy is "rolled out" (executed) and the supervisor provides demonstrations for all  new states, which are then aggregated with the previous examples for the next iteration. DAgger and related algorithms have been applied in a wide range of applications, from quadrotor flight to natural language to Atari games~\cite{NIPS2014_5421,duvallet2013imitation,ross2013learning}. It has been shown for DAgger, the number of errors to scale linearly for strongly convex loss functions.

One drawback is that such online algorithms increase the burden on the supervisor. Our goal is to reduce supervisor burden without degrading robot performance, by only requesting demonstrations for risky states.

Risk arises because:  (1) states are different than the states previously encountered and used to generate the current policy, i.e., they can be considered novel or outliers~\cite{hodge2004survey}; or (2) states are in a region that have been misclassified, in previous training data.
%\adnote{Ken's version below does not seem to represent what we actually do.}
%the action associated with a state under the current policy yield a resulting state that is inconsistent with the expected next state. [how do we measure 2? - Ken]

We introduce the SHIV algorithm (Svm-based reduction in Human InterVention), which can reduce supervisor burden by only requesting supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary state distributions.. It builds on the One Class SVM method for approximating quantile level sets \cite{scholkopf2001estimating} to provide a classification method for deciding if a state is risky or safe.

@@

\noindent\textbf{Active Learning from Demonstration}
Online learning from demonstration can be formulated as a stream-based active learning problem ~\cite{atlas1990training,cohn1994improving}. In stream based active learning, the decision of whether to query the supervisor (for a control signal or not) is not over the entire state space (like in traditional pool-based active learning), but only on the states encountered during learning.

@@

In SHIV, we use a query selection method that allows for a non-stationary distribution, evaluating risk based on novelty (is the state an outlier to the training distribution), as in Fig. \ref{fig:support_example}(c). Additionally, we also assign higher risk to  historically mislabeled regions, as in Fig. \ref{fig:support_example}(d). We then use a classification method based on the One Class SVM \cite{scholkopf2001estimating} to predict if a state is risky or not risky, which we we define as safe.%Our approach builds on the One Class SVM method for approximating quantile level sets \cite{}, which has already been used in novelty detection on high-dimensional image data \cite{}.

\noindent\textbf{Risk via Novelty Detection.}
One part of our risk definition is the notion that a trained model will be able to generalize within the distribution it is
trained on \cite{tokdar2010importance}. Novelty detection \cite{hodge2004survey} is concerned with recognizing that a sample is outside of this distribution.
One approach to novelty detection is to directly estimate the underlying probability density function from which the data is sampled. If the probability of that point is low with respect to the underlying density, it will be marked as novel. However, the amount of data needed to accurately do so scales exponentially in the number of dimensions \cite{nadaraya1964estimating}.

@@

Our approach is based on the One Class SVM proposed by Scholk{\"o}pf et al., which estimates a particular quantile level set for the training data by solving a convex quadratic program  to find support
vectors~\cite{scholkopf2001estimating}. The method has been theoretically shown to approximate the quantile levelset of a density estimate
asymptotically for correctly chosen bandwidth settings and in the case of a normalized Gaussian kernel function \cite{vert2006consistency}. 
In \cite{liu2014unsupervised}, the One Class SVM has furthermore been used for novelty detection in high-dimensional image data. To our knowledge the use of One Class SVM to online learning from demonstration has never been done before. 
\begin{figure*}[t]
\centering

@@

\section{Problem Statement}
The goal of this work is to 1)learn a policy, such that for states visited after completion of training the difference in controls taken by the robot is as small as possible to that of the supervisor 2)the supervisor is only requested for a query in states, which the robot is mis-predicting the supervisor's control.

\noindent\textbf{Modeling Choices and Assumptions} We model the system dynamics as markovian and stochastic. The system dynamics are model to be stationary (i.e. given a state and a control the probability of the next state doesn't change over time). We model the initial state as sampled from a distribution over the state space.

We assume a known state space and set of controls. We assume access to a robot or simulator, such that we can start in some initial state and apply a series of controls and observe the following states that occur.  Lastly, we assume access to a supervisor that can provide a series of demonstrated trajectories and can be given a state visited by the robot or simulator and provide a control signal of what the supervisor would do.

\noindent\textbf{Policies and State Densities.}
We denote by $\mathcal{X}$ the set consisting of observable states for a robot task, consisting, for example, of \mlnote{We should discuss if we are going to consider images states or not}
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of allowable control inputs for the robot, which can be discrete or
continuous. We assume Markovian dynamics, such that the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$: 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume a probability density over initial states $p(\bx_0)$.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

@@

Let $p(\bx_t|\theta)$ denote the value of the density of states visited at time $t$ if the robot follows the policy
$\pi_{\theta}$ from time $0$ to time $t-1$. This density can be computed by marginalization: $p(\bx_t|\theta) =
\int_{\bx_{t-1}}...\int_{\bx_1} p((\bx_t,...,\bx_1)|\theta) d\bx_{t-1}...d\bx_1$. Following \cite{ross2010reduction}, we can compute
the average density on states for any timepoint by 
\begin{equation}
p(\bx|\theta) = \frac{1}{T} \sum^T_{t=1} p(\bx_t|\theta)
\label{eq:density}
\end{equation}

@@

\noindent\textbf{Objective.} The objective of policy learning is to find a policy that minimizes some known cost function
$C(\hat{\tau}) = \sum^T_{t=1} c(\bx_t,\bu_t)$ of a trajectory $\hat{\tau}$. The cost function $c:\mathcal{X}\times \mathcal{U}\to \mathbb{R}$ is typically user defined and task specific. 
For example, in task of inserting a peg into a hole, a function on distance between the peg's current and desired final state can
be considered \cite{levine2015end}.  
In our learning from demonstration setup, the robot does not have access to the cost function itself. Instead, it only has access to 
a `supervisor' that we assume uses some $\tilde{\pi}$ to minimize $C(\hat{\tau})$, to an acceptable level. \mlnote{the last statement is intended to mean implementors of this algorithm have agreed that the supervisor may not be optimal, but has a level of performance that is worth automating} We are furthermore given
an initial set of $N$ stochastic demonstration trajectories $\lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying this policy. This induces a training dataset $\mathcal{D}$ of all state-control input pairs from the demonstrated trajectories.

@@

Given a candidate policy $\pi_{\theta}$, we then use the surrogate loss function to approximately measure how "close" the policy's
returned control input $\pi_{\theta}(\bx)\in \mathcal{U}$ at a given state $\bx\in \mathcal{X}$ is to the supervisor's policy's control output
$\tilde{\pi}(\bx)\in \mathcal{U}$. The goal is to produce a that minimizes the surrgoate loss between relative to the supervisor's policy.

@@

 \vspace{-2ex}
\begin{align}\label{eq:LFD_obj}
\underset{\theta}{\min} \: E_{p(\bx|\theta)} [l(\pi_\theta(\bx),\tilde{\pi}(\bx))].
\end{align}
with $p(\bx|\theta)$ from Eq. \ref{eq:density}.

 If the robot could learn the policy  perfectly, this state density would match the one encountered in the user examples. But if the robot to makes an error, it changes the distribution of states over which it visits, which in practice can lead it to encounter states in which it has no labels from the expert and can't generalize too \cite{pomerleau1989alvinn}. Thus, the motivation for iterative algorithms like DAgger and now SHIV, which iterate between learning a policy and then the supervisor providing feedback. The feedback is in the form of control signals on states sampled from the robot's new distribution of states.

@@

Both SHIV and DAgger \cite{ross2010reduction} solve the minimization in Eq. \ref{eq:LFD_obj} by iterating two steps: 1) compute a $\theta$ using the training data $\mathcal{D}$ thus far, and 2) execute the policy induced by the current $\theta$, and ask for labels for the encountered states. However, instead of querying the supervisor for every new state, SHIV actively decides whether the state is risky to warrant a query.

\subsection{Step 1.}
The first step of any iteration $k$ is to compute a $\theta_k$ that minimizes surrogate loss on the current dataset $\mathcal{D_k}=\{(x_i,u_i)|i\in\{1,\ldots,M\}\}$ of demonstrated state-control pairs (initially just the set $D$ from above):

@@

This problem is a supervised learning problem that can be solved using machine learning techniques, for example a support vector machine or a neural network. Finding the right function representation and optimization method for a given problem is still an open question in machine learning, however a large number of advances have been made that make the use of these techniques very feasible \cite{scholkopf2002learning}

To handle the fact that the supervisor's policy can be noisy, a zero-mean noise term $\epsilon$ 
can be considered as present in policy's output.  A regularization technique in the optimization is
used to control the smoothness of the function that is fit to the sampled data. In practice this regularization corresponds to a penalty term on either the L2 norm on the weights for regression based techniques or the slack coefficient for support vector machines \cite{scholkopf2002learning}.
 \subsection{Step 2}
The second step starts in both SHIV and DAgger by rolling out the policy $\pi_{\theta_{k}}$ to sample states that are likely to occur under $\theta_{k}$.

@@

Dagger has been experimentally shown to perform well in both simulation and real world
    experiments~\cite{NIPS2014_5421,ross2010reduction,ross2013learning}. Under a `no regret' assumption, Ross et al.\cite{ross2010reduction} furthermore
deduced error bounds on the expected loss, which scale linearly in the time horizon $T$ --- as opposed to offline methods which scale quadratically. For a conservative $\sigma$, the SHIV performance is analogous to DAgger without asking the expert when it is absolutely not necessary. However, if $\sigma$ is too high, then SHIV will not collect new data that is needed and revert back to the performance of offline methods. We analyze the dependence on $\sigma$ in Sec. \ref{sec:car}. Our results suggest that SHIV is robust to this parameter selection.

@@

\vspace{-2ex}
\begin{align}\label{eq:decision_func}
g(\bx) = \mbox{sgn}(\langle w,\Phi(x) \rangle-\rho).
\end{align}

@@

Our experiments analyze the performance of risk as a query selection strategy, with our without the addition of the second risk criteria --- regions of space where the surrogate loss is high.

@@

\section{Experiments}
All experiments were run on a machine with OS X with a 2.7 GHz Intel core i7 processor and 16 GB
1600 MHz memory in Python 2.7. The policies, $\pi_\theta$ were represented using either the linear SVM or  kernelized ridge regression classes in Scikit-Learn~\cite{scikit-learn}.

>>>>

\newboolean{include-notes}
\setboolean{include-notes}{true}
\newcommand{\adnote}[1]{\ifthenelse{\boolean{include-notes}}%
 {\textcolor{blue}{\textbf{AD: #1}}}{}}

@@

\author{Michael Laskey$^1$, Sam Staszak$^3$, Wesley Yu-Shu Hsieh$^1$, Jeff Mahler$^1$, \\ Florian T. Pokorny$^1$, Anca D. Dragan$^1$, and Ken Goldberg$^{1,2}$% <-this % stops a space
\thanks{$^1$ Department of Electrical Engineering and Computer Sciences; {\small \{mdlaskey,iamwesleyhsieh, ftpokorny, anca\}@berkeley.edu}}%
\thanks{$^2$ Department of Industrial Engineering and Operations Research; {\small goldberg@berkeley.edu}}%
\thanks{$^{1-2}$ University of California, Berkeley;  Berkeley, CA 94720, USA}%
}
\begin{document}

@@

\begin{abstract}
Online algorithms for robot model-free\adnote{model-free robot} learning from demonstration (such as DAgger) can learn policies for problems where \adnote{the} system
dynamics and \adnote{the} cost function are unknown. However, during learning they impose a burden on supervisors to provide labels (control
signals) as the robot encounters new states when iteratively executing its current best policy. We introduce the SHIV
algorithm (Svm-based reduction in Human InterVention), which reduces this burden by enabling the robot to \emph{only request supervision for risky states}. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary high dimensional state distributions.  To facilitate scaling and outlier rejection, risk is based on distance to an approximate level set boundary defined by a One Class support vector machine.  We compare SHIV with DAgger in three policy learning experiments: 1) a driving simulator with a 27936D visual HOG feature state representation, 2) 4DOF robot arm push-grasping in clutter (simulation with Box2D) with a 22D state representation  of joint angles and the pose of each object, and 3) surgical needle insertion with a 16D state represented spatial transform.  Results suggest that SHIV can reduce the number of queries up to 70$\%$\adnote{double check this}.

@@

In model-free robot learning from demonstration, a robot learns to perform a task, such as driving or grasping an object in a cluttered environment (Fig. \ref{fig:teaser}), from examples provided by a competent\adnote{why competent? called expert in the literature; either stick to that or simplify to just supervisor} supervisor, usually a human. The problem is model-free if  the system dynamics are unknown, e.g. dynamics model may be highly non-linear or noisy.\adnote{the problem isn't model free, the approach is; when the dynamics are not known, you can explicitly learn them, or learn a policy; that's the difference between model based and model free} In learning from demonstration, the cost function, which is the negative of the reward function, is usually unknown or significantly delayed\adnote{confusing; in learning from demonstration it is always unknown. unclear what significantly delayed means and it feels like you are confusing reinforcement with imitation learning}. In the grasping for clutter example, shown in Fig. \ref{fig:teaser}(b)\adnote{missing comma} the robot is only penalized when an object is knocked off the table, which means a potentially large number of states can be visited before any feedback.\adnote{this is a reinforcement learning problem then, not a learning from demonstration problem} To potentially\adnote{potentially? sounds like a filler word} avoid this a supervisor provides guidance through demonstrations and the robot learns a policy mapping states to controls~\cite{argall2009survey}.  \adnote{Maybe keep first sentence and replace the rest with: In such problems, the robot does not have access to neither the cost function that it should optimize, nor the dynamics model. The former happens when it is difficult for a human to specify how to trade-off various aspects that matter, like [fill in and refer to driving]. The latter happens when either the system of the interaction with the world is difficulty to characterize, like when the robot is pushing obstacles out of the way while reaching for a goal object (Fig. \ref{fig:teaser}(b). Rather than learning the cost function and the dynamics model, and then using optimization to produce a policy for the task, the robot learns the policy directly from supervisor examples mapping states to controls~\cite{argall2009survey}.}

In the supervised\adnote{all of them are the supervised setting} setting, a policy is \adnote{avoid passive voice} learned offline from a fixed set of examples.\adnote{In the offline setting, the robot learns the policy based on a batch of examples, and then executes it to achieve the task.} When the robot executes this policy\adnote{During execution,}, the sequential process\adnote{what sequential process? you haven't said anything about sequential yet. During execution, a small error can accumulate and push the robot away from the region of the state space where it has seen examples.} can quickly evolve away from the previously explored set of states.  The learned policy might not be helpful in unfamiliar states, leading to unrecoverable failures.  For example, a robot driving may be trained on examples driving safely down the center of a lane, but even slight deviations will put the robot into states near the side of the road where it has not encountered before, where its policy could  fail~\cite{pomerleau1989alvinn}. It has been shown\adnote{by jesus}, the number of errors made by the robot in such cases  can scale quadratically with the time horizon of the task~\cite{ross2010efficient}.\adnote{avoid passive, replace with "The number of errors made by the robot can, in the worst case, scale linearly..."}

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/teaser.pdf}
\caption{ 
\todo{update figure}\adnote{make everything flush in a grid, add caption to graph, increase font size in graph labels so that it is the same as the caption}
SHIV \\adnote[(orange in the graph in d)} reduces the number of queries to the supervisor \adnote{compared to the state of the art?} in our three test domains: 1) a driving simulator where the goal is to learn a controller on HOG features extracted from synthetic images to keep a car on a polygonal track; 2) push-grasping in clutter in a physics simulator, where the goal is to learn a controller from human demonstrations to grasp the green object while not letting the red square objects fall off the gray boundary representing a table; 3) learning a controller for the Da Vinci Research Kit from a master surgeon's demonstrations of correcting a needle before suturing in surgery. 
}
\vspace*{-10pt}
\label{fig:teaser}
\end{figure}

On-line\adnote{online;it's how you spell it in the rest of the document, and you don't spell offline off-line, so I don't get the reason for this inconsistency} learning from demonstration approaches such as DAgger \adnote{address this issue by }iteratively gather more examples from the supervisor~\cite{grollman2007dogged,ross2010efficient,ross2010reduction}. In DAgger, a series of policies are learned\adnote{DAgger learns a series of policies}.  In each iteration, the latest policy is "rolled out" (executed) and the supervisor provides demonstrations for all  new states, which are then aggregated with the previous examples for the next iteration\adnote{the learning step is missing here; At each itteration, the robot trains a policy based on the existing examples, then rolls out (executes) that policy and ....}. DAgger and related algorithms have been applied in a wide range of applications, from quadrotor flight to natural language to Atari games~\cite{NIPS2014_5421,duvallet2013imitation,ross2013learning}. It has been shown for DAgger, the number of errors to scale linearly for strongly convex loss functions.\adnote{clunky sentence. how about In DAgger, the number of errors scales only linearly with the time horizon of the task. -- also too early to talk about loss functions; note also the "only", because we use this to contrast with the end of the last paragraph, where it was quadratically}

One drawback is that\adnote{perhaps too weak of a statement; maybe "To achieve better performance guarantees, online algorithms singificatonly increase the burden on the supervisor, who now iteratively labels all states that the robot visits during training.} such online algorithms increase the burden on the supervisor. Our goal is to reduce supervisor burden without degrading robot performance,\adnote{performance.} \adnote{check spelling, there are many typos in this text}
\begin{quote}
\adnote{\emph{Our key insight is that the robot only needs to request demonstrations for risky states.}}
\end{quote}

Risk arises because:  (1) states are different than the states previously encountered and used to generate the current policy, i.e., they can be considered novel or outliers~\cite{hodge2004survey}; or (2) states are in a region that have\adnote{grammar:the region has, not have} been misclassified, in previous training data.
%\adnote{Ken's version below does not seem to represent what we actually do.}
%the action associated with a state under the current policy yield a resulting state that is inconsistent with the expected next state. [how do we measure 2? - Ken]

\adnote{We contribute an algorithm for online learning from demonstration that can actively decide whether it is necessary to ask the supervisor to label the new states that the robot encounters. Our algorithm, SHIV (...), reduced supervisor burden by..}
We introduce the SHIV algorithm (Svm-based reduction in Human InterVention), which can reduce supervisor burden by only requesting supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to \adnote{the} \emph{non-stationary} state distributions \adnote{the robot encounters as it iteratively executes learned policies}. It\adnote{I think we should say we here: we build on..} builds on the One Class SVM method for approximating quantile level sets \cite{scholkopf2001estimating} to provide a classification method for deciding if a state is risky or safe.

\adnote{we need results here; both the highlights of the analysis we do (so that people don;t think we are blindly comparing it to dagger without asking if all these changes are necessary), as well as the main point, that shiv is better for the same budget, and refer to fig 1}

@@

\noindent\textbf{Active Learning from Demonstration}
Online learning from demonstration can be formulated as a stream-based active learning problem ~\cite{atlas1990training,cohn1994improving}. In stream based active learning, the decision of whether to query the supervisor (for a control signal or not)\adnote{(for a label or control signal) or not -- note also where the ) ends} is not over the entire state space (like in traditional pool-based active learning), but only on the states encountered during learning. \adnote{I think streambased is more general then that, where the stream can come from some oracle and not from any learning process.}

@@

In SHIV, we use a query selection method that allows for a non-stationary distribution, evaluating risk based on novelty (is the state an outlier to the training distribution), as in Fig. \ref{fig:support_example}(c). Additionally, we also assign higher risk to  historically mislabeled regions, as in Fig. \ref{fig:support_example}(d). We then use a classification method based on the One Class SVM \cite{scholkopf2001estimating} to predict if a state is risky or not risky, which we we define as safe.%Our approach builds on the One Class SVM method for approximating quantile level sets \cite{}, which has already been used in novelty detection on high-dimensional image data \cite{}. 
\adnote{we need more here on why this is good for a nonstationary distribution; michael, give it a try and send me a paragraph via email, i;ll help edit it}

\noindent\textbf{Risk via Novelty Detection.}
One part of our risk definition is the notion that a trained model will be able to generalize within the distribution it is
trained on \cite{tokdar2010importance}\adnote{,making states outside of this distribution risky}. Novelty detection \cite{hodge2004survey} is concerned with recognizing \adnote{when this happens: recognizing }that a sample is outside of this distribution.

One approach to novelty detection is to directly estimate the underlying probability density function from which the data is sampled. If the probability of that point is low with respect to the underlying density, it will be marked as novel. However, the amount of data needed to accurately do so scales exponentially in the number of dimensions \cite{nadaraya1964estimating}.

@@

Our approach is based on the One Class SVM proposed by Scholk{\"o}pf et al., which estimates a particular quantile level set for the training data by solving a convex quadratic program  to find support
vectors~\cite{scholkopf2001estimating}. The method has been theoretically shown to approximate the quantile levelset of a density estimate
asymptotically for correctly chosen bandwidth settings and in the case of a normalized Gaussian kernel function \cite{vert2006consistency}. 
In \cite{liu2014unsupervised}, the One Class SVM has furthermore been used for novelty detection in high-dimensional image data. To our knowledge the use of One Class SVM to online learning from demonstration has never been done before. \adnote{you can say this to me, but not really write it in a paper; I think that perhaps you want to say, and in a new paragraph so that it stands out, "Our contribution is the addition of stream-based active learning to online learning from demonstration. To do so, we use a query selection method based on the One Class SVM.}

\begin{figure*}[t]
\centering

@@

\section{Problem Statement}
The goal of this work is to 1) learn a policy, such that for states visited after completion of training the difference in controls taken by the robot is as small as possible to that of the supervisor\adnote{very complicated, could we simplify to "learn a policy that matches that of the supervisor's?} 2)the supervisor is only requested for a query in states, which the robot is mis-predicting the supervisor's control.\adnote{gramma weird here; the robot only asks the supervisor for an example for states in which it would have mis-predicted the supervisor's control.; but this is a bit strong, as we don't achieve this goal; maybe we want to learn a policy that matches that of the supervisor's while asking the supervisor for as few examples as possible?}

\noindent\textbf{Modeling Choices and Assumptions} We model the system dynamics as markovian and stochastic. The system dynamics are model\adnote{first, you mean modeled; second, by jesus? :-) if it's we, say we. We model the system dynamics ad markovian, stochastic, and stationary.} to be stationary (i.e. given a state and a control the probability of the next state doesn't change over time). We model the initial state as sampled from a distribution over the state space.

We assume a known state space and set of controls. We assume access to a robot or simulator\adnote{simulator when you don't knowo the dynamics is weird; maybe say access to a robot (or a computatonally expensive simulator)}, such that we can start in some initial state and apply a series of controls and observe the following states that occur.\adnote{too coloquial; that can sample from the state sequences induced by a sequence of controls.} \adnote{I feel like this is a very weak assumption and we should not even bother mentioning that there is a robot -- of course there is, otherwise how are you even getting demonstrations.. hmm.. and it shows up later as well} Lastly, we assume access to a supervisor that can provide a series of demonstrated trajectories and can be given a state visited by the robot or simulator and provide a control signal of what the supervisor would do. \adnote{clunky; Lastly, we assume access to a supervisor who can, given a state, provide the desired control signal label. (note that doing this also means you can do trajectories, so that's redundant}

\adnote{weird that this subsection comes before some of the math, I keep wanting to refer to the math symbols and write the right equations for these things, like markovian and probability over the initial state and so on..}

\noindent\textbf{Policies and State Densities.}
We denote by $\mathcal{X}$ the set consisting of observable states for a robot task, consisting, for example, of \mlnote{We should discuss if we are going to consider images states or not}
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of allowable control inputs for the robot, which can be discrete or
continuous. We assume Markovian dynamics\adnote{see, this shows up again, and much better here when I know what the notation is and I can write it in math as well}, such that the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$: 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume a probability density over initial states $p(\bx_0)$.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

@@

Let $p(\bx_t|\theta)$ denote the value of the density of states visited at time $t$ if the robot follows the policy
$\pi_{\theta}$ from time $0$ to time $t-1$. This density can be computed by marginalization: $p(\bx_t|\theta) =
\int_{\bx_{t-1}}...\int_{\bx_1} p((\bx_t,...,\bx_1)|\theta) d\bx_{t-1}...d\bx_1$. Following \cite{ross2010reduction}, we can compute
the average density on states \emph{for any timepoint} by 
\begin{equation}
p(\bx|\theta) = \frac{1}{T} \sum^T_{t=1} p(\bx_t|\theta)
\label{eq:density}
\end{equation}

@@

\noindent\textbf{Objective.} The objective of \adnote{typical (you want to not make the person think that this is our objective too)} policy learning is to find a policy that minimizes some known cost function
$C(\hat{\tau}) = \sum^T_{t=1} c(\bx_t,\bu_t)$ of a trajectory $\hat{\tau}$. The cost function $c:\mathcal{X}\times \mathcal{U}\to \mathbb{R}$ is typically user defined and task specific. 
For example, in task of inserting a peg into a hole, a function on distance between the peg's current and desired final state can
be considered \cite{levine2015end}.

In our learning from demonstration setup, \adnote{however,}the robot does not have access to the cost function itself. Instead, it only has access to 
a `supervisor'\adnote{why quotes all of the sudden?} that we assume uses some $\tilde{\pi}$ to minimize $C(\hat{\tau})$, to an acceptable level. \mlnote{the last statement is intended to mean implementors of this algorithm have agreed that the supervisor may not be optimal, but has a level of performance that is worth automating} We are furthermore given
an initial set of $N$ stochastic demonstration trajectories $\lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying this policy. This induces a training dataset $\mathcal{D}$ of all state-control input pairs from the demonstrated trajectories.

@@

Given a candidate policy $\pi_{\theta}$, we then use the surrogate loss function to approximately measure how "close" the policy's
returned control input $\pi_{\theta}(\bx)\in \mathcal{U}$ at a given state $\bx\in \mathcal{X}$ is to the supervisor's policy's control output
$\tilde{\pi}(\bx)\in \mathcal{U}$. The goal is to produce a \adnote{policy}that minimizes the surrgoate loss between relative to the supervisor's policy.

@@

 \vspace{-2ex}
\begin{align}\label{eq:LFD_obj}
\underset{\theta}{\min} \: E_{p(\bx|\theta)} [l(\pi_\theta(\bx),\tilde{\pi}(\bx))]
\end{align}
with $p(\bx|\theta)$ from Eq. \ref{eq:density}.

 If the robot could learn the policy  perfectly, this state density would match the one encountered in the user examples. But if the robot to\adnote{delete to} makes an error, it changes the distribution of states over\adnote{delete over..} which it visits, which in practice can lead it to encounter states in which it has no labels from the expert and can't generalize too \cite{pomerleau1989alvinn}.\adnote{clunky: But if the robot makes an error, that error changes the distribution of states that the robot will visit, which can lead to states that are far away from any examples and difficult to generalize to.} Thus,\adnote{this sentence has no verb. maybe "this is the motivation.."} the motivation for iterative algorithms like DAgger and now SHIV, which iterate between learning a policy and then the supervisor providing feedback. The feedback is in the form of control signals on states sampled from the robot's new distribution of states.

@@

Both SHIV and DAgger \cite{ross2010reduction} solve the minimization in Eq. \ref{eq:LFD_obj} by iterating two steps: 1) compute a $\theta$ using the training data $\mathcal{D}$ thus far, and 2) execute the policy induced by the current $\theta$, and ask for labels for the encountered states. However, instead of querying the supervisor for every new state, SHIV actively decides whether the state is risky \adnote{enough }to warrant a query.

\subsection{Step 1.}
The first step of any iteration $k$ is to compute a $\theta_k$ that minimizes surrogate loss on the current dataset $\mathcal{D}_k=\{(x_i,u_i)|i\in\{1,\ldots,M\}\}$ of demonstrated state-control pairs (initially just the set $\mathcal{D}$ of initial trajectory demonstrations):

@@

This problem is a supervised learning problem that can be solved using machine learning techniques\adnote{this sentence makes use seem like we don't know machine learning; it's subtle but could hurt us; how about this is a standard supervised learning problem, solvable by estimators like a support vector machine or a neural net.}, for example a support vector machine or a neural network. Finding the right function representation and optimization method for a given problem is still an open question in machine learning, however a large number of advances have been made that make the use of these techniques very feasible \cite{scholkopf2002learning}\adnote{is this really needed? what does it add? it basically says figuring out what method to use is hard; maybe say: performance will vary with the selection of a particular estimator \cite{}}

To handle the fact that the supervisor's policy can be noisy, a zero-mean noise term $\epsilon$ 
can be considered as present in \adnote{the} policy's output\adnote{does espilon show up anywhere? do we need to say this? maybe a footnote instead?}.  A regularization technique in the optimization is
used to control the smoothness of the function that is fit to the sampled data. In practice this regularization corresponds to a penalty term on either the L2 norm on the weights for regression based techniques or the slack coefficient for support vector machines \cite{scholkopf2002learning}.
 \subsection{Step 2}
The second step starts in both SHIV and DAgger by rolling out the policy $\pi_{\theta_{k}}$ to sample states that are likely to occur under $\theta_{k}$.

@@

Dagger has been experimentally shown to perform well in both simulation and real world
    experiments~\cite{NIPS2014_5421,ross2010reduction,ross2013learning}. Under a `no regret' assumption, Ross et al.\cite{ross2010reduction} furthermore
deduced error bounds on the expected loss, which scale linearly in the time horizon $T$ --- as opposed to offline methods which scale quadratically. For a conservative $\sigma$, the SHIV performance is analogous to DAgger without asking the expert when it is absolutely not necessary. However, if $\sigma$ is too high, then SHIV will not collect new data that is needed and revert back to the performance of offline methods. We analyze the dependence on $\sigma$ in Sec. \ref{sec:car}, and our results suggest that SHIV is robust to this parameter selection.

@@

\vspace{-2ex}
\begin{align}\label{eq:decision_func}
g(\bx) = \mbox{sgn}(\langle w,\Phi(x) \rangle-\rho).
\end{align}
\adnote{we need $\sigma$ showing up here as $g_{\sigma}$ and remind the reader that the that inner product is a function of $\sigma$. might also be good to add a sentence or two on the intuitive role of sigma in this decision -- it caputures how much we expect states to generalize, i.e. how smooth the policy function is}

@@

Our experiments analyze the performance of risk as a query selection strategy, with or without the addition of the second risk criteria, i.e. regions of space where the surrogate loss is high.

@@

\section{Experiments}
